---
title:  "0. A Guide to readJDX Vignettes"
author:
  - name: Bryan A. Hanson^1^
    email: hanson@depauw.edu
    affiliation: |
        1. Professor Emeritus of Chemistry & Biochemistry, DePauw University, Greencastle IN USA.
date:  "`r Sys.Date()`"
output:
    bookdown::html_document2: # use for pkgdown site
    #bookdown::pdf_document2: # use for CRAN to keep size down; breaks GHA
      toc: no
      toc_depth: 3
      fig_caption: yes
      number_sections: true
# Required: Vignette metadata for inclusion in a package.
vignette: >
    %\VignetteIndexEntry{0. Guide to readJDX Vignettes}
    %\VignetteKeyword{JCAMP-DX, NMR, IR, Raman, MS, UV-Vis}
    %\VignettePackage{readJDX}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
link-citations: yes
bibliography: JCAMP-DX.bib
#biblio-style: plain
pkgdown:
  as_is: true
---

```{r results = "hide", echo = FALSE}
library("readJDX")
rm(list = ls())
desc <- packageDescription("readJDX")
```

This vignette is based on `readJDX` version `r desc$Version`.

---

The following vignettes are available, in addition to this one:

* *readJDX*, a general introduction and overview.
* The following vignettes will be useful for those wishing to contribute to the development of `readJDX` and those that are troubleshooting files:
    + *Taxonomy* describes a taxonomy of JCAMP-DX file formats along with a diagram of which `readJDX` functions call each other.
    + *MiniDIFDUP_1* is an annotated version of the `MiniDIFDUP` data set illustrating the ASDF compression scheme.
    + *MiniDIFDUP_2* is an expanded and annotated version of the `MiniDIFDUP` data set.

To see the last two vignettes (in PDF format), do the following:

```{r, eval = FALSE}
browseVignettes("readJDX")
```
