---
title: Program Flow in readJDX

# Use letters for affiliations
author:
  - name: Bryan A. Hanson
    affiliation: a

address:
  - code: a
    address: Dept. of Chemistry \& Biochemistry, DePauw University; \url{hanson@depauw.edu}

# Place DOI URL or CRAN Package URL here
doi: "github.com/bryanhanson/readJDX"

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 10pt

# Optional: Force one-column layout, default is two-column
one_column: true

# Optional: Enable section numbering, default is unnumbered
#numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Bibliography
#bibliography: JCAMP-DX

# Optional: Enable a 'Draft' watermark on the document
watermark: false

# Customize footer, eg by referencing the vignette
footer_contents: "Program Flow"

# Produce a pinp document
output: pinp::pinp

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{Program Flow}
  %\VignetteKeyword{JCAMP, DX, NMR, IR, Raman}
  %\VignettePackage{readJDX}
  %\VignetteEngine{knitr::rmarkdown}

header-includes:
  - \usepackage{booktabs}
  - \usepackage{colortbl}
---

```{r results = "hide", echo = FALSE}
library("readJDX")
suppressPackageStartupMessages(library("kableExtra"))
rm(list = ls())

desc <- packageDescription("readJDX")
set.seed(999)
knitr::opts_chunk$set(tidy = TRUE,
  fig.height = 7, fig.width = 7,
  fig.fullwidth = TRUE, echo = FALSE)
```

This vignette is based on `readJDX` version `r desc$Version`.

# Program Flow

`readJDX` is coded in such a way that it should be easy to add features.  Contributions to improve or expand the package, including pull requests, are always welcome!  Table \ref{tab:PF} shows the overall flow of the function calls.  Only a couple of these functions are exported, so take a look at the source code for documentation.  Be sure to check out the *MiniDIFDUP_1* and *MiniDIFDUP_2* vignettes for additional information about the JCAMP-DX file structure and how `readJDX` functions extract the data.

```{r progFlow, results = "asis"}
fName <- c(
  "readJDX",
  "- findVariableLists",
  "- extractParams",
  "- processVariableList",
  "- - processXYY", # 5
  "- - - decompLines",
  "- - - - getComp",
  "- - - - unSQZ",
  "- - - - insertDUPs",
  "- - - - - repDUPs", # 10
  "- - - - deDIF",
  "- - - - - unDIF",
  "- - - - yValueCheck",
  "- - processPT"
  )

fInput <- c(
  "file name",
  "character vector: all lines from original file",
  "character vector: just the metadata lines",
  "character vector: a single VL",
  "character vector: a single VL", # 5
  "character vector: a single VL",
  "character vector: a single VL",
  "character vector: named with the ASDF mode, called via lapply",
  "character vector: named with the ASDF mode, called via lapply",
  "a string of length one", # 10
  "*list* of character vectors from a single VL; the character vectors are named with the ASDF mode, the list is named with line numbers",
  "character vector: one line from VL, entries named by ASDF code",
  "*list* of character vectors from a single VL; the character vectors are named with the ASDF mode, the list is named with line numbers",
  "character vector: a single VL"
  )

DF <- data.frame(fName, fInput)
names(DF) <- c("function", "input")

kable(DF, "latex", booktabs = TRUE, caption = "Program Flow.\\label{tab:PF}") %>% kable_styling(latex_options = c("striped", "hold_position")) %>% footnote(general = c("VL stands for variable list, as defined in the JCAMP-DX standard. For examples see the *MiniDIFDUP_1* vignette."))
```
